---
title: "Getting Started with stargazer2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with stargazer2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "",
  message  = FALSE,
  warning  = FALSE
)
library(stargazer2)
has_wooldridge <- requireNamespace("wooldridge",  quietly = TRUE)
has_sandwich   <- requireNamespace("sandwich",    quietly = TRUE)
```

`stargazer2` is a drop-in replacement for the `stargazer` package with native support for modern econometrics packages.
For `lm` objects the output is designed to be identical to the original, with one key addition: the standard error type is always identified in the table note.

## Dataset

We use the `wage1` dataset from the `wooldridge` package.
Three categorical variables are constructed from existing binary indicators for use as controls and clustering variables.

```{r wage1-setup, eval = requireNamespace("wooldridge", quietly=TRUE)}
library(wooldridge)
data(wage1)

wage1$region <- factor(
  ifelse(wage1$northcen == 1, "northcen",
  ifelse(wage1$south    == 1, "south",
  ifelse(wage1$west     == 1, "west", "northeast"))),
  levels = c("northeast", "northcen", "south", "west")
)

wage1$occupation <- factor(
  ifelse(wage1$profocc == 1, "professional",
  ifelse(wage1$clerocc == 1, "clerical",
  ifelse(wage1$servocc == 1, "service", "other"))),
  levels = c("other", "professional", "clerical", "service")
)

wage1$industry <- factor(
  ifelse(wage1$construc == 1, "construction",
  ifelse(wage1$ndurman  == 1, "nondurable_manuf",
  ifelse(wage1$trcommpu == 1, "transport",
  ifelse(wage1$trade    == 1, "trade",
  ifelse(wage1$services == 1, "services",
  ifelse(wage1$profserv == 1, "prof_services", "other")))))),
  levels = c("other", "construction", "nondurable_manuf",
             "transport", "trade", "services", "prof_services")
)
```

## A familiar table

Four progressively richer log-wage specifications:

```{r lm-models, eval = requireNamespace("wooldridge", quietly=TRUE)}
m1 <- lm(lwage ~ educ + exper + tenure, wage1)
m2 <- lm(lwage ~ educ + exper + tenure + female + married, wage1)
m3 <- lm(lwage ~ educ + exper + tenure + female + married +
            region + occupation, wage1)
m4 <- lm(lwage ~ educ + exper + tenure + female + married +
            region + occupation + industry, wage1)
```

A single call produces a publication-ready table.
Models 3 and 4 include factor variables; `omit` suppresses their level dummies from the coefficient block so the table stays focused on the economic variables of interest.

```{r basic-table, eval = requireNamespace("wooldridge", quietly=TRUE), results = "asis"}
stargazer(m1, m2, m3, m4,
          type             = "html",
          title            = "Determinants of Log Wages",
          dep.var.labels   = "log(Wage)",
          covariate.labels = c("Education", "Experience", "Tenure",
                               "Female", "Married"),
          omit             = c("region", "occupation", "industry"),
          column.labels    = c("Baseline", "Demographics",
                               "Region/Occ.", "Full"),
          notes.append     = FALSE,
          notes            = "Controls for region, occupation, and industry included in models (3) and (4).")
```

## Output formats

The `type` argument selects the output format.
The default is `"latex"` for embedding in papers; `"text"` provides a quick terminal preview; `"html"` renders tables directly in R Markdown documents like this one.

### Text (ASCII)

```{r text-output, eval = requireNamespace("wooldridge", quietly=TRUE)}
stargazer(m1, m2,
          type             = "text",
          dep.var.labels   = "log(Wage)",
          covariate.labels = c("Education", "Experience", "Tenure"))
```

### LaTeX source

```{r latex-output, eval = requireNamespace("wooldridge", quietly=TRUE)}
stargazer(m1, m2,
          type             = "latex",
          title            = "Determinants of Log Wages",
          label            = "tab:wage-ols",
          dep.var.labels   = "log(Wage)",
          covariate.labels = c("Education", "Experience", "Tenure"))
```

Paste this into your `.tex` file or write it directly with `out = "table.tex"`.

## Custom standard errors via `vcov`

The `vcov` argument accepts a list of variance-covariance matrices — one per model.
`stargazer2` extracts the square root of the diagonal internally and updates the table note to name the SE type used in each column.
This works with any function that returns a matrix: `sandwich::vcovHC`, `sandwich::vcovCL`, `lmtest`-style wrappers, or your own estimator.

### HC1-robust SEs

```{r robust-se, eval = requireNamespace("wooldridge", quietly=TRUE) && requireNamespace("sandwich", quietly=TRUE), results = "asis"}
library(sandwich)
stargazer(m1, m2, m3, m4,
          type             = "html",
          dep.var.labels   = "log(Wage)",
          covariate.labels = c("Education", "Experience", "Tenure",
                               "Female", "Married"),
          omit             = c("region", "occupation", "industry"),
          column.labels    = c("Baseline", "Demographics",
                               "Region/Occ.", "Full"),
          vcov             = list(vcovHC(m1, type = "HC1"),
                                  vcovHC(m2, type = "HC1"),
                                  vcovHC(m3, type = "HC1"),
                                  vcovHC(m4, type = "HC1")))
```

### Industry-clustered SEs

```{r clustered-se, eval = requireNamespace("wooldridge", quietly=TRUE) && requireNamespace("sandwich", quietly=TRUE), results = "asis"}
stargazer(m1, m2, m3, m4,
          type             = "html",
          dep.var.labels   = "log(Wage)",
          covariate.labels = c("Education", "Experience", "Tenure",
                               "Female", "Married"),
          omit             = c("region", "occupation", "industry"),
          column.labels    = c("Baseline", "Demographics",
                               "Region/Occ.", "Full"),
          vcov             = list(vcovCL(m1, cluster = ~industry, data = wage1),
                                  vcovCL(m2, cluster = ~industry, data = wage1),
                                  vcovCL(m3, cluster = ~industry, data = wage1),
                                  vcovCL(m4, cluster = ~industry, data = wage1)))
```

### Mixed SE types across columns

`vcov` entries need not be the same type across columns.
When SE types differ, the note reports them by column group — here column (1) uses HC1-robust SEs while columns (2)–(4) use industry-clustered SEs.

```{r mixed-se, eval = requireNamespace("wooldridge", quietly=TRUE) && requireNamespace("sandwich", quietly=TRUE), results = "asis"}
stargazer(m1, m2, m3, m4,
          type             = "html",
          dep.var.labels   = "log(Wage)",
          covariate.labels = c("Education", "Experience", "Tenure",
                               "Female", "Married"),
          omit             = c("region", "occupation", "industry"),
          column.labels    = c("Baseline", "Demographics",
                               "Region/Occ.", "Full"),
          vcov             = list(vcovHC(m1, type = "HC1"),
                                  vcovCL(m2, cluster = ~industry, data = wage1),
                                  vcovCL(m3, cluster = ~industry, data = wage1),
                                  vcovCL(m4, cluster = ~industry, data = wage1)))
```

## Cosmetic options

The most commonly used formatting arguments are illustrated here together.

| Argument | Purpose |
|---|---|
| `dep.var.labels` | Override dependent variable name(s) |
| `covariate.labels` | Rename coefficient rows (in display order) |
| `column.labels` | Column headers beneath the dep-var line |
| `omit` / `keep` | Regex patterns to drop or retain coefficient rows |
| `digits` | Decimal places for all numbers |
| `star.cutoffs` | P-value thresholds for significance stars |
| `notes` / `notes.append` | Add or replace the automatic table note |
| `title` / `label` | Caption and `\label{}` for LaTeX |

## Table styles

The `style` argument selects a layout preset.

| Style | Layout | Significance note |
|---|---|---|
| `"stargazer2"` | Single `\hline`, full-width left-aligned note | p-value thresholds (default) |
| `"stargazer"` | Matches original package exactly (double rules, `\\[-1.8ex]`) | p-value thresholds |
| `"aer"` | American Economic Review — clean, no dep-var caption | Text descriptions ("Significant at the X percent level") |
| `"qje"` | Quarterly Journal of Economics — like AER; observations labelled $N$ | Text descriptions |

The difference is most visible in the LaTeX source.
Here is a two-column table with the `"aer"` style:

```{r style-aer, eval = requireNamespace("wooldridge", quietly=TRUE)}
stargazer(m1, m2,
          type             = "latex",
          dep.var.labels   = "log(Wage)",
          covariate.labels = c("Education", "Experience", "Tenure"),
          style            = "aer")
```

## Summary statistics

Passing a data frame instead of model objects produces a summary statistics table.

```{r summary-stats, eval = requireNamespace("wooldridge", quietly=TRUE)}
stargazer(
  wage1[, c("lwage", "educ", "exper", "tenure", "female", "married")],
  type             = "text",
  covariate.labels = c("log(Wage)", "Education", "Experience",
                       "Tenure", "Female", "Married")
)
```
