---
title: "Panel Data Models with plm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Panel Data Models with plm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "",
  message  = FALSE,
  warning  = FALSE
)
library(stargazer2)
```

`stargazer2` supports `plm` model objects natively. It auto-detects the
estimator type (FE, RE, Pooled OLS, FD, Between), displays fixed-effect and
random-effect indicator rows, and reports the appropriate fit statistics from
`summary.plm`.

## Dataset and models

The `Grunfeld` dataset (10 US manufacturing firms, 1935--1954, balanced panel)
is included with `plm`. We estimate four specifications of an investment
equation:

```{r models, eval = requireNamespace("plm", quietly = TRUE)}
library(plm)
data("Grunfeld", package = "plm")

m_pool <- plm(inv ~ value + capital, Grunfeld,
              index = c("firm", "year"), model = "pooling")

m_fe   <- plm(inv ~ value + capital, Grunfeld,
              index = c("firm", "year"), model = "within")

m_twfe <- plm(inv ~ value + capital, Grunfeld,
              index = c("firm", "year"), model = "within", effect = "twoways")

m_re   <- plm(inv ~ value + capital, Grunfeld,
              index = c("firm", "year"), model = "random")
```

## Default output

A bare call produces a complete table. `stargazer2` reads the model type from
each `plm` object and builds indicator rows for the fixed and random effects
present across all columns:

```{r default-text, eval = requireNamespace("plm", quietly = TRUE)}
stargazer(m_pool, m_fe, m_twfe, m_re, type = "text")
```

Things to notice:

- **Model-type row** identifies each column as Pooled OLS, FE, or RE.
- **Firm FE** and **Year FE** rows track which specifications absorb each
  set of fixed effects. Both show "Yes" only in the two-way FE column.
- **Firm RE** row shows "Yes" for the random-effects column and "No"
  elsewhere.
- The **Constant** disappears from the FE columns — the within transformation
  absorbs it.
- The RE column's F statistic is a Wald chi-square test and is shown without
  degrees-of-freedom annotation.

## Robust standard errors

`plm` ships its own vcov functions. Pass them through the `vcov` argument
exactly as with `lm` or `fixest` models. `vcovHC(..., method = "arellano")`
gives standard errors clustered by the individual unit, the most common choice
for FE models in applied work:

```{r robust-text, eval = requireNamespace("plm", quietly = TRUE)}
V_fe   <- vcovHC(m_fe,   method = "arellano")
V_twfe <- vcovHC(m_twfe, method = "arellano")

stargazer(m_fe, m_twfe,
          type     = "text",
          vcov     = list(V_fe, V_twfe),
          se_label = "Arellano cluster-robust standard errors")
```

Because `plm::vcovHC` returns a plain matrix with no class attribute,
`stargazer2` cannot auto-detect the SE type. Supply `se_label =` explicitly
for a precise table note, as shown above.

## Adding formatting options

Once the defaults look right, labels can be added. The LaTeX source below
sets a title and cross-reference label, renames the dependent variable and
covariates, and assigns custom column headers:

```{r formatted-latex, eval = requireNamespace("plm", quietly = TRUE)}
stargazer(m_pool, m_fe, m_twfe, m_re,
          type             = "latex",
          title            = "Investment Equations: Grunfeld Panel Data",
          label            = "tab:grunfeld",
          dep.var.labels   = "Investment",
          covariate.labels = c("Market Value", "Capital Stock"),
          column.labels    = c("Pooled OLS", "FE", "Two-way FE", "RE"))
```
